Rheology of Cubic Blue Phases

### 2013-06-05 papers/bprheo/bprheo.tex
Update bprheo paper
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig/trunk@1999 40876441-3496-0410-a613-a2fca35b2ae2

%% ****** Start of file apstemplate.tex ****** %
%%
%%
%%   This file is part of the APS files in the REVTeX 4 distribution.
%%   Version 4.1r of REVTeX, August 2010
%%
%%
%%   Copyright (c) 2001, 2009, 2010 The American Physical Society.
%%
%%   See the REVTeX 4 README file for restrictions and more information.
%%
%
% This is a template for producing manuscripts for use with REVTEX 4.0
% Copy this file to another name and then work on that file.
% That way, you always have this original template file to use.
%
% Group addresses by affiliation; use superscriptaddress for long
% author lists, or if there are many overlapping affiliations.
% For Phys. Rev. appearance, change preprint to twocolumn.
% Choose pra, prb, prc, prd, pre, prl, prstab, prstper, or rmp for journal
%  Add 'draft' option to mark overfull boxes with black boxes
%  Add 'showpacs' option to make PACS codes appear
%  Add 'showkeys' option to make keywords appear
\documentclass[aps,pre,reprint,superscriptaddress, twocolumn]{revtex4}
%\documentclass[aps,pre,preprint,superscriptaddress]{revtex4}
%\documentclass[aps,prl,reprint,groupedaddress]{revtex4}

% You should use BibTeX and apsrev.bst for references
% Choosing a journal automatically selects the correct APS
% BibTeX style file (bst file), so only uncomment the line
% below if necessary.
\bibliographystyle{apsrev4-1}

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{float}

\newcommand{\e}[1]{\times10^{#1}}
\newcommand{\gd}{\dot{\gamma}}
\newcommand{\com}[1]{{\bf {#1}}}

@@

% Use the \preprint command to place your local institutional report
% number in the upper righthand corner of the title page in preprint mode.
% Multiple \preprint commands are allowed.
% Use the 'preprintnumbers' class option to override journal defaults
% to display numbers if necessary
%\preprint{}

%Title of paper
\title{Rheology of Cubic Blue Phases}

% repeat the \author .. \affiliation  etc. as needed
% \email, \thanks, \homepage, \altaffiliation all apply to the current
% author. Explanatory text should go in the []'s, actual e-mail
% address or url should go in the {}'s for \email and \homepage.
% Please use the appropriate macro foreach each type of information

% \affiliation command applies to all authors since the last
% \affiliation command. The \affiliation command should follow the
% other information
% \affiliation can be followed by \email, \homepage, \thanks as well.

\author{O. Henrich}
\email[Corresponding author: ]{ohenrich@epcc.ed.ac.uk}
\affiliation{EPCC, School of Physics and Astronomy, University of Edinburgh,\\JCMB Kings Buildings, Mayfield Road, Edinburgh EH9 3JZ, UK}
\affiliation{Centre for Computational Science, University College London,\\20 Gordon Street, London WC1H 0AJ, UK}

\author{K. Stratford}
\affiliation{EPCC, School of Physics and Astronomy, University of Edinburgh,\\JCMB Kings Buildings, Mayfield Road, Edinburgh EH9 3JZ, UK}

\author{P.V. Coveney}
\affiliation{Centre for Computational Science, University College London,\\20 Gordon Street, London WC1H 0AJ, UK}

\author{M.E. Cates}
\affiliation{SUPA, School of Physics and Astronomy, University of Edinburgh,\\JCMB Kings Buildings, Mayfield Road, Edinburgh EH9 3JZ, UK}

\author{D. Marenduzzo}
\affiliation{SUPA, School of Physics and Astronomy, University of Edinburgh,\\JCMB Kings Buildings, Mayfield Road, Edinburgh EH9 3JZ, UK}

%\homepage[]{Your web page}
%\thanks{}
%\altaffiliation{}

%Collaboration name if desired (requires use of superscriptaddress
%option in \documentclass). \noaffiliation is required (may also be
%used with the \author command).
%\collaboration can be followed by \email, \homepage, \thanks as well.
%\collaboration{}

%\noaffiliation

\date{\today}

\begin{abstract}
We study the behaviour of cubic blue phases under shear flow
via lattice Boltzmann simulations. We focus on the two experimentally observed
phases, Blue Phase I (BPI) and Blue Phase II (BPII). The disclination network of Blue Phase 
II continuously breaks and reforms under shear, leading to an oscillatory stress response in time.
For larger shear rates, the network breaks up into a cholesteric helix, lying either along the
vorticity or the flow gradient direction according to the imposed flow.
BPI leads to a very different response. Here, oscillations are only possible for intermediate 
shear rates -- very slow flow leads to a transition into an amorphous network 
with a residual stress. The flow-induced oscillation regime features a novel arrangement
of double helical disclinations which rotate due to the shear. Large flow can, 
just as in the BPII case, lead to the rupture of the network and the stabilisation
of less ordered cholesteric structures. Our results provide the first 
theoretical investigation of sheared blue phases in large systems, which
are relevant to understanding the bulk rheology of these materials. 
\end{abstract}

% insert suggested PACS numbers in braces on next line
\pacs{47.11.Qr, 47.57.Lj, 47.57.Qk, 61.30.Dk, 61.30.Pq, 61.30.Jf, 83.80.Xz}
% insert suggested keywords - APS authors don't need to do this
\keywords{blue phase, cholesterics, chiral nematics, rheology}

%\maketitle must follow title, authors, abstract, \pacs, and \keywords
\maketitle

% body of paper here - Use proper section commands
% References should be done using the \cite, \ref, and \label commands
\section{Introduction}
% Put \label in argument of \section for cross-referencing
%\section{\label{}}

Cholesterics are liquid crystals in which the local nematic director field spontaneously 
twists in thermodynamic equilibrium~\cite{deGennes}. The preferred configuration close 
to the isotropic boundary features twist around two perpendicular axes, as opposed to 
just one axis in the regular cholesteric state, and the corresponding deformation is 
denoted a ``double-twist cylinder''.
As it is topologically impossible to cover continuously 3D space with double-twist 
cylinders, defects arise. These organise into a variety of regular periodic lattices, 
giving rise to the so-called cubic blue phases (BPs)~\cite{Grebel:1984,Wright:1989}. 
There are two experimentally observed cubic blue phases, BPI and BPII (a third, BPIII, 
is thought to be amorphous~\cite{Henrich:2011a}).

BPs have been long considered as purely of academic interest due to their very narrow 
range of stability. This view has changed since the creation of polymer-stabilised and 
temperature-stabilised BPs~\cite{Kikuchi:2002,Coles:2005}, which has opened up the 
possibility of novel applications.
During the last few years considerable progress has been achieved regarding the behaviour 
of BPs in confined geometries~\cite{Fukuda:2010a, Fukuda:2010b, Ravnik:2011b}, under 
external fields ~\cite{Alexander:2008,Fukuda:2009,Henrich:2010a,Castles:2010,Tiribocchi:2011a}, 
and in the presence of colloidal particles~\cite{Ravnik:2011a}.
The kinetics of BP domain growth have been recently addressed in~\cite{Henrich:2010b}. 
However, our understanding of their dynamical behaviour under flow remains
very limited. The aim of this work is to address this issue by studying,
for the first time, the response of confined BP samples to a shear flow.

Flow response in cholesterics is both strongly non-Newtonian and highly anisotropic.
For example, if a cholesteric helix is subjected to a Poiseuille flow along
the helical axis, small pressure differences drive flow mainly through
''permeation'', first investigated by Helfrich~\cite{Helfrich:1969}.
In the permeation mode the liquid crystal flows while leaving the director
field virtually unchanged, which leads to high dissipation and large
viscosities. Marenduzzo et al.~\cite{Marenduzzo:2006a,Marenduzzo:2006b} simulated 
shear and Poiseuille flow in cholesteric liquid crystals in the permeation mode, and 
showed the importance of the boundary conditions in determining the apparent viscosity of the fluid. 
They also found that a strong secondary flow appears.
Rey~\cite{Rey:1996a, Rey:1996b} studied shear in cholesterics oriented with the helix along 
the vorticity axis and found that, at low Ericksen number, travelling twist waves appear which 
lead to the rotation of the cholesteric helix. At higher forcing, the helix uncoils and 
leaves a flow-induced nematic phase.
Rey also studied cholesterics subjected to both steady flow and low frequency
small amplitude oscillatory shear for different helix orientations
~\cite{Rey:2000, Rey:2002}. He found that splay/bend/twist deformations were
excited when the helix was aligned along the flow direction; splay/bend
deformation occurred when the helix was aligned along the velocity gradient;
but only twist deformations
appeared when the helix was aligned along the vorticity axis.

Dupuis et al.~\cite{Dupuis:2005} performed the first numerical investigation of BP rheology 
in Poiseuille flow, starting from equilibrium structures of BPI and BPII and a periodic 
array of doubly twisted cylinders.
Under small forcing, the network opposed the flow giving rise to a significant 
increase in apparent viscosity.
Upon increasing the forcing they found clear evidence of shear thinning.
In the crossover region they predicted a novel oscillatory regime where the network 
continuously breaks and reforms as portions of the disclinations in the centre of the channel 
move to neighbouring cells and relink with the parts of the network left behind by the flow. 
Compared with the cholesteric case, the viscosity still decreases with forcing 
(the system shear thins) but much less than
for cholesterics in the permeation mode, which is in agreement with experiments
~\cite{Zapotocky:1999, Ramos:2002}.

Our work differs from these earlier efforts as it addresses shear flows, 
and focuses on flow-induced reconstruction and nonequilibrium transition
between different blue phase networks, which appear at intermediate or
high shear. Our simulations employ Lees-Edwards boundary conditions, which
are naturally suited to address these regimes, and bulk as opposed
to boundary-dominated flow. 
Our simulations are large scale and parallel, so that we can study
significantly larger systems than previously possible, so that our
results can be tested with bulk rheology experiments.

Our paper is organised as follows. 
In Section II, we describe the method we use and review
the hydrodynamic equation of motion which we aim to solve. In Section III, 
we report our numerical results, separating them into subsections relative to
blue phase I and blue phase II, and corresponding to low, intermediate, and
high shear. Finally, we draw our conclusions in Section IV.

@@

Our approach is based on the well-established Beris-Edwards model for hydrodynamics of
cholesteric liquid crystals \cite{Beris:1994}, which describes the ordered state 
in terms of a traceless, symmetric tensor order parameter ${\mathbf Q}({\mathbf r})$. 
In the uniaxial approximation, the order parameter is given by
$Q_{\alpha \beta}= q_s ( \hat{n}_\alpha \hat{n}_\beta - \frac{1}{3}\; \delta_{\alpha\beta})$
with $\hat{{\mathbf n}}$ the director field and $q_s$ the amplitude of nematic
order. More generally,
the largest eigenvalue of ${\mathbf Q}$, $0\le q_s\le\frac{2}{3}$
characterises the local degree of orientational order.
The thermodynamic properties of the liquid crystal are determined by a free energy
${\cal F}$, whose density $f$ consists of a bulk contribution $f_b$ and a gradient part $f_g$, as follows,
\begin{eqnarray}
f_b&=&\frac{A_0}{2}\left(1-\frac{\gamma}{3}\right) Q_{\alpha \beta}^2\nonumber\\
&-&\frac{A_0 \gamma}{3}Q_{\alpha \beta} Q_{\beta \gamma} Q_{\gamma \alpha}+\frac{A_0 \gamma}{4}(Q_{\alpha \beta}^2)^2,\nonumber\\
f_g&=&\frac{K}{2}(\varepsilon_{\alpha\gamma\delta} \partial_\gamma Q_{\delta\beta}+2 q_0 Q_{\alpha \beta})^2+\frac{K}{2}(\partial_\beta Q_{\alpha \beta})^2.\label{FE}
\end{eqnarray}
The first term contains the bulk-free energy constant $A_0$ and the inverse temperature $\gamma$ which controls the magnitude of order.
The second part quantifies the cost of elastic distortions, which are proportional to the elastic constant $K$;
we work for simplicity in the one-elastic constant approximation. The wavevector $q_0$ is equal to $2\pi/p_0$, where $p_0$ is the cholesteric pitch.
The actual periodicity of the BP structure, $p$, does not need to be equal to $p_0$.
The ``redshift'' $r=p/p_0$ is adjusted during the simulation, to optimise the free energy density
at all times -- this is done by following the procedure
previously described in~\cite{Alexander:2006}.

A thermodynamic state is specified by two dimensionless quantities: the effective temperature 
\begin{equation}
\tau=\frac{27(1-\gamma/3)}{\gamma}
\end{equation}
and chirality 
\begin{equation}
\kappa=\sqrt{\frac{108 K q_0^2}{A_0 \gamma}},
\end{equation}
which measures the ratio of gradient to bulk free energy.

The dynamical evolution of the order parameter is given by the equation 
\begin{equation}
\left(\partial_t+ v_\alpha \partial_\alpha \right){\mathbf Q} - {\mathbf S}({\mathbf W},{\mathbf Q}) = \Gamma {\mathbf H}.
\label{op-eom}
\end{equation}
The first term on the left hand side of Eq.\ref{op-eom} is a material derivative, which describes the rate of change of a quantity advected by the flow.
The second term accounts for the rate of change due to local velocity gradients $W_{\alpha \beta}=\partial_\beta v_\alpha$,
and is explicitly given by
\begin{eqnarray}
{\mathbf S}({\mathbf W}, {\mathbf Q}) &=& (\xi {\mathbf A} + {\boldsymbol \Omega})({\mathbf Q}+\frac{\mathbf I}{3})\nonumber\\
& &\hspace*{-1.5cm}+ ({\mathbf Q}+\frac{\mathbf I}{3})(\xi {\mathbf A}  - {\boldsymbol \Omega})-2 \xi ({\mathbf Q}+\frac{\mathbf I}{3})
\mathrm{Tr}({\mathbf Q W}),
\label{sw}
\end{eqnarray}
where $\mathrm{Tr}$ denotes the tensorial trace, while 
${\mathbf A}=({\mathbf W}+{\mathbf W}^T)/2$ and
${\boldsymbol \Omega}=({\mathbf W}-{\mathbf W}^T)/2$ are the symmetric and antisymmetric part of the velocity gradient, respectively. $\xi$ 
is a constant depending on the molecular details of the liquid crystal.
Flow alignment occurs if $\xi \cos{2\theta}=(3q_s)/(2+q_s)$ has a real solution, where $\theta$ is the 
co-called Leslie angle: we select this case by 
setting $\xi=0.7$ in our simulations.
${\mathbf H}$ is the molecular field, which is the functional derivative of $\cal F$ that preserves the tracelessness of $\mathbf Q$:
\begin{equation}
{\bf H}=-\frac{\delta {\cal F}}{\delta {\bf Q}}+\frac{\bf I}{3}\,
\mathrm{Tr} \left(\frac{\delta {\cal F}}{\delta {\bf Q}}\right).
\label{molfield}
\end{equation}
The rotational diffusion constant $\Gamma$ in Eq.~\ref{op-eom} is proportional
to the inverse of the rotational viscosity $\gamma_1=2 q_s^2/\Gamma$
\cite{deGennes}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%This is the article LaTeX template for RSC journals
%Copyright The Royal Society of Chemistry 2010
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[8.5pt,twoside,twocolumn]{article}
\oddsidemargin -1.2cm
\evensidemargin -1.2cm
\textwidth 18cm
\headheight 1.0in
\topmargin -3.5cm
\textheight 22cm
\usepackage[super,sort&compress,comma]{natbib} 
\usepackage{mhchem}
\usepackage{times,mathptmx}
%\usepackage{times}
% feel free not to use mathptmx if it causes difficulties
\usepackage{sectsty}
\usepackage{balance} 
\usepackage{color}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{float}
\newcommand{\e}[1]{\times10^{#1}}
\newcommand{\gd}{\dot{\gamma}}
\newcommand{\com}[1]{{\bf {#1}}}
\newcommand{\rev}[1]{{\textcolor{red}{#1}}}

\usepackage{graphicx} %eps figures can be used instead
\usepackage{lastpage}
\usepackage[format=plain,justification=raggedright,singlelinecheck=false,font=small,labelfont=bf,labelsep=space]{caption} 
\usepackage{fancyhdr}
\pagestyle{fancy}

\newif\ifold
%\oldtrue

@@

\thispagestyle{plain}
\fancypagestyle{plain}{
\fancyhead[L]{\includegraphics[height=8pt]{headers/LH}}
\fancyhead[C]{\hspace{-1cm}\includegraphics[height=20pt]{headers/CH}}
\fancyhead[R]{\includegraphics[height=10pt]{headers/RH}\vspace{-0.2cm}}
\renewcommand{\headrulewidth}{1pt}}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand\footnoterule{\vspace*{1pt}% 
\hrule width 3.4in height 0.4pt \vspace*{5pt}} 
\setcounter{secnumdepth}{5}

\makeatletter 
\def\subsubsection{\@startsection{subsubsection}{3}{10pt}{-1.25ex plus -1ex minus -.1ex}{0ex plus 0ex}{\normalsize\bf}} 
\def\paragraph{\@startsection{paragraph}{4}{10pt}{-1.25ex plus -1ex minus -.1ex}{0ex plus 0ex}{\normalsize\textit}} 
\renewcommand\@biblabel[1]{#1}            
\renewcommand\@makefntext[1]% 
{\noindent\makebox[0pt][r]{\@thefnmark\,}#1}
\makeatother 
\renewcommand{\figurename}{\small{Fig.}~}
\sectionfont{\large}
\subsectionfont{\normalsize}

\fancyfoot{}
\fancyfoot[LO,RE]{\vspace{-7pt}\includegraphics[height=9pt]{headers/LF}}
\fancyfoot[CO]{\vspace{-7.2pt}\hspace{12.2cm}\includegraphics{headers/RF}}
\fancyfoot[CE]{\vspace{-7.5pt}\hspace{-13.5cm}\includegraphics{headers/RF}}
\fancyfoot[RO]{\footnotesize{\sffamily{1--\pageref{LastPage} ~\textbar  \hspace{2pt}\thepage}}}
\fancyfoot[LE]{\footnotesize{\sffamily{\thepage~\textbar\hspace{3.45cm} 1--\pageref{LastPage}}}}
\fancyhead{}
\renewcommand{\headrulewidth}{1pt} 
\renewcommand{\footrulewidth}{1pt}
\setlength{\arrayrulewidth}{1pt}
\setlength{\columnsep}{6.5mm}
\setlength\bibsep{1pt}

\twocolumn[
  \begin{@twocolumnfalse}
\noindent\LARGE{\textbf{Rheology of Cubic Blue Phases}}
\vspace{0.6cm}

\noindent\large{\textbf{Oliver Henrich $^{\ast}$\textit{$^{a,b}$}, Kevin Stratford \textit{$^{a}$}, Peter V. Coveney \textit{$^{b}$}, Michael E. Cates \textit{$^{c}$}, Davide Marenduzzo \textit{$^{c}$}}}\vspace{0.5cm}
%Please note that \ast indicates the corresponding author(s) but no footnote text is required.

\noindent\textit{\small{\textbf{Received Xth XXXXXXXXXX 20XX, Accepted Xth XXXXXXXXX 20XX\newline
First published on the web Xth XXXXXXXXXX 200X}}}

\noindent \textbf{\small{DOI: }}
\vspace{0.6cm}
%Please do not change this text.

\noindent \normalsize{We study the behaviour of cubic blue phases under shear flow 
via lattice Boltzmann simulations. We focus on the two experimentally observed
phases, Blue Phase I (BPI) and Blue Phase II (BPII). The disclination network of Blue Phase 
II continuously breaks and reforms under steady shear, leading to an oscillatory stress response in time.
For larger shear rates, the structure breaks up \rev{into a Grandjean texture with a 
cholesteric helix lying along the flow gradient direction according to the imposed flow}.
BPI leads to a very different response. Here, oscillations are only possible for intermediate
shear rates -- very slow flow leads to a transition into an amorphous network 
with an apparent yield stress. \rev{Larger shear rates can make the network sensitive
to flow-induced transitions to metastable and amorphous states. For even
larger flow rates a Grandjean texture is observed, just as in the case of BPII.
At the highest flow rates both cubic blue phases adopt a flow-aligned nematic state.}
Our results provide the first theoretical investigation of sheared blue phases in large systems, and are relevant to understanding the bulk rheology of these materials.}
\vspace{0.5cm}
 \end{@twocolumnfalse}
  ]

%\footnotetext{\dag~Electronic Supplementary Information (ESI) available: [details of any supplementary information available should be included here]. See DOI: 10.1039/b000000x/}

%Please use \dag to cite the ESI in the main text of the article.
%If you article does not have ESI please remove the the \dag symbol from the title and the above footnotetext.

\footnotetext{\textit{$^{a}$~EPCC, School of Physics and Astronomy, University of Edinburgh,\\JCMB Kings Buildings, Mayfield Road, Edinburgh EH9 3JZ, UK; E-mail: ohenrich@epcc.ed.ac.uk}}
\footnotetext{\textit{$^{b}$~Centre for Computational Science, University College London,\\20 Gordon Street, London WC1H 0AJ, UK}}
\footnotetext{\textit{$^{c}$~SUPA, School of Physics and Astronomy, University of Edinburgh,\\JCMB Kings Buildings, Mayfield Road, Edinburgh EH9 3JZ, UK}}

%additional addresses can be cited as above using the lower-case letters, c, d, e... If all authors are from the same address, no letter is required

%\footnotetext{\ddag~Additional footnotes to the title and authors can be included \emph{e.g.}\ `Present address:' or `These authors contributed equally to this work' as above using the symbols: \ddag, \textsection, and \P. Please place the appropriate symbol next to the author's name and include a \texttt{\textbackslash footnotetext} entry in the the correct place in the list.}

\section{Introduction}
% Put \label in argument of \section for cross-referencing
%\section{\label{}}

Cholesterics are liquid crystals in which the local nematic director field shows spontaneous 
twist in thermodynamic equilibrium~\cite{deGennes}. 
The simplest manifestation is the standard cholesteric phase where the
director precesses around a single helical axis of fixed orientation.
For highly chiral systems, however,
the preferred configuration close 
to the isotropic boundary features twist around two perpendicular axes, as opposed to 
just one axis in the regular cholesteric state, and the corresponding deformation is 
denoted a ``double-twist cylinder''.
As it is topologically impossible to cover continuously 3D space with double-twist 
cylinders, defects arise. The resulting disclination lines (at which the
nematic director is undefined)
 organise into a variety of regular periodic lattices, 
giving rise to the so-called cubic blue phases (BPs)~\cite{Grebel:1984,Wright:1989}. 
There are two experimentally observed cubic blue phases, BPI and BPII (a third, BPIII, 
is thought to be amorphous~\cite{Henrich:2011a}).

BPs were long considered as purely of academic interest due to their very narrow 
range of stability. This view has changed since the creation of polymer-stabilised and other thermally 
stabilised BPs~\cite{Kikuchi:2002,Coles:2005}, which has opened up the 
possibility of novel applications.
During the last few years considerable progress has been achieved regarding the behaviour 
of BPs in confined geometries~\cite{Fukuda:2010a, Fukuda:2010b, Ravnik:2011b}, under 
external fields~\cite{Alexander:2008,Fukuda:2009,Henrich:2010a,Castles:2010,Tiribocchi:2011a}, 
and in the presence of colloidal particles~\cite{Ravnik:2011a}.
The kinetics of BP domain growth have been recently addressed~\cite{Henrich:2010b}. 
However, our understanding of their dynamical behaviour under flow remains
very limited. The aim of this work is to address this issue by studying,
for the first time, the response of large BP samples to a shear flow.

Flow response in cholesterics is both strongly non-Newtonian and highly anisotropic.
For example, if a standard cholesteric phase is subjected to a Poiseuille flow along
its helical axis, small pressure differences drive flow mainly through
``permeation'', as first investigated by Helfrich~\cite{Helfrich:1969}.
In the permeation mode the liquid crystal flows while leaving the director
field virtually unchanged, which leads to high dissipation and large
viscosities. Marenduzzo et al.~\cite{Marenduzzo:2006a,Marenduzzo:2006b} simulated 
shear and Poiseuille flow in cholesteric liquid crystals in the permeation mode, and 
showed the importance of the boundary conditions in determining the apparent viscosity of the fluid. 
They also found that a strong secondary flow appears.
\rev{Early experiments with cholesteric liquid crystals
showed that flow can also give rise to conformational transitions. \cite{Press:1978}.}
Rey~\cite{Rey:1996a, Rey:1996b} studied shear in cholesterics oriented with the helix along 
the vorticity axis and found that, at low Ericksen number, travelling twist waves appear which 
lead to the rotation of the cholesteric helix. At higher forcing, the helix uncoils, creating a flow-induced nematic phase. \rev{These result, however, were derived under the 
assumption that the molecules rotate only in flow-gradient plane whilst
the orientation of the cholesteric helix remained unchanged.}  
Rey also studied cholesterics subjected to both steady flow and low frequency
small amplitude oscillatory shear for different helix orientations
~\cite{Rey:2000, Rey:2002}. He found that splay/bend/twist deformations were
excited when the helix was aligned along the flow direction; splay/bend
deformation occurred when the helix was aligned along the velocity gradient;
but only twist deformations
appeared when the helix was aligned along the vorticity axis.

Dupuis et al.~\cite{Dupuis:2005} performed the first numerical investigation of BP rheology 
in Poiseuille flow, starting from equilibrium structures of BPI and BPII and a periodic 
array of doubly twisted cylinders.
Under small forcing, the network opposed the flow, giving rise to a significant 
increase in apparent viscosity.
Upon increasing the forcing they found clear evidence of shear thinning.
In the crossover region they predicted a novel oscillatory regime where the network 
continuously breaks and reforms as portions of the disclinations in the centre of the channel 
move to neighbouring cells and relink with the parts of the network left behind by the flow. 
The viscosity still decreases with forcing 
(the system shear thins) but much less than
for cholesterics in the permeation mode, which is in agreement with experiments
~\cite{Zapotocky:1999, Ramos:2002}.

Our work differs from these earlier efforts as it addresses homogeneous shear flows, 
and focuses on flow-induced reconstruction and nonequilibrium transition
between different blue phase networks, which appear at intermediate or
high shear. Our simulations employ Lees-Edwards boundary conditions, which
are naturally suited to address these regimes, and bulk as opposed
to boundary-dominated flow. 
Our simulations are large scale and parallel, so that we can study
significantly larger systems than previously possible, and our
results can in principle be compared with bulk rheology experiments.

Our paper is organised as follows. 
In Section II, we describe the method we use and review
the hydrodynamic equations of motion which we aim to solve. In Section III, 
we report our numerical results, separating them into subsections referring to
Blue Phase I and Blue Phase II, and corresponding to low, intermediate, and
high shear. Finally, we draw our conclusions in Section IV.

@@

Our approach is based on the well-established Beris-Edwards model for hydrodynamics of
cholesteric liquid crystals \cite{Beris:1994}, which describes the ordered state 
in terms of a traceless, symmetric tensor order parameter ${\mathbf Q}({\mathbf r})$. 
In the uniaxial approximation, the order parameter is given by
$Q_{\alpha \beta}= q_s ( \hat{n}_\alpha \hat{n}_\beta - \frac{1}{3}\; \delta_{\alpha\beta})$
with $\hat{{\mathbf n}}$ the director field and $q_s$ the amplitude of nematic
order. More generally,
the largest eigenvalue of ${\mathbf Q}$, $0\le q_s\le\frac{2}{3}$
characterises the local degree of orientational order.
The thermodynamic properties of the liquid crystal are determined by a free energy
${\cal F}$, whose density $f$ consists of a bulk contribution $f_b$ and a gradient part $f_g$, as follows,
\begin{eqnarray}
f_b&=&\frac{A_0}{2}\left(1-\frac{\gamma}{3}\right) Q_{\alpha \beta}^2\nonumber\\
&-&\frac{A_0 \gamma}{3}Q_{\alpha \beta} Q_{\beta \gamma} Q_{\gamma \alpha}+\frac{A_0 \gamma}{4}(Q_{\alpha \beta}^2)^2,\nonumber\\
f_g&=&\frac{K}{2}(\varepsilon_{\alpha\gamma\delta} \partial_\gamma Q_{\delta\beta}+2 q_0 Q_{\alpha \beta})^2+\frac{K}{2}(\partial_\beta Q_{\alpha \beta})^2.\label{FE}
\end{eqnarray}
The first term contains a bulk-free energy constant $A_0$ and the temperature-related parameter $\gamma$ which controls the magnitude of order.
The second part quantifies the cost of elastic distortions, which is proportional to the elastic constant $K$;
we work for simplicity in the one-elastic constant approximation~\cite{deGennes}. The wavevector $q_0$ is equal to $2\pi/p_0$, where $p_0$ is the cholesteric pitch.
The actual periodicity of the BP structure, $p$, does not need to be equal to $p_0$.
Indeed, the ``redshift'' $r=p/p_0$ is adjusted during the equilibration phase of the 
simulation, to optimise the free energy density
before shearing begins -- this is done by following the procedure
previously described in~\cite{Alexander:2006}.

A thermodynamic state is specified by two dimensionless quantities: the reduced temperature 
\begin{equation}
\tau=\frac{27(1-\gamma/3)}{\gamma},
\end{equation}
which vanishes at the spinodal point of a nematic ($q_0=0$), 
and the reduced chirality 
\begin{equation}
\kappa=\sqrt{\frac{108 K q_0^2}{A_0 \gamma}},
\end{equation}
which measures the ratio of gradient to bulk free energy.

The dynamical evolution of the order parameter is given by the equation 
\begin{equation}
\left(\partial_t+ v_\alpha \partial_\alpha \right){\mathbf Q} - {\mathbf S}({\mathbf W},{\mathbf Q}) = \Gamma {\mathbf H}.
\label{op-eom}
\end{equation}
The first term on the left hand side of Eq.\ref{op-eom} is a material derivative, which describes the rate of change of a quantity advected by the flow.
The second term accounts for the rate of change due to local velocity gradients $W_{\alpha \beta}=\partial_\beta v_\alpha$,
and is explicitly given by~\cite{Beris:1994}
\begin{eqnarray}
{\mathbf S}({\mathbf W}, {\mathbf Q}) &=& (\xi {\mathbf A} + {\boldsymbol \Omega})({\mathbf Q}+\frac{\mathbf I}{3})\nonumber\\
& &\hspace*{-1.5cm}+ ({\mathbf Q}+\frac{\mathbf I}{3})(\xi {\mathbf A}  - {\boldsymbol \Omega})-2 \xi ({\mathbf Q}+\frac{\mathbf I}{3})
\mathrm{Tr}({\mathbf Q W}),
\label{sw}
\end{eqnarray}
where $\mathrm{Tr}$ denotes the tensorial trace, while 
${\mathbf A}=({\mathbf W}+{\mathbf W}^T)/2$ and
${\boldsymbol \Omega}=({\mathbf W}-{\mathbf W}^T)/2$ are the symmetric and antisymmetric part of the velocity gradient, respectively. $\xi$ 
is a constant depending on the molecular details of the liquid crystal.
Flow alignment occurs if $\xi \cos{2\theta}=(3q_s)/(2+q_s)$ has a real solution for $\theta$, the 
so-called Leslie angle: we select this case by 
setting $\xi=0.7$ in our simulations.
${\mathbf H}$ is the molecular field, which is a functional derivative of $\cal F$ that respects the tracelessness of $\mathbf Q$:
\begin{equation}
{\bf H}=-\frac{\delta {\cal F}}{\delta {\bf Q}}+\frac{\bf I}{3}\,
\mathrm{Tr} \left(\frac{\delta {\cal F}}{\delta {\bf Q}}\right).
\label{molfield}
\end{equation}
The rotational diffusion constant $\Gamma$ in Eq.~\ref{op-eom} is proportional
to the inverse of the rotational viscosity $\gamma_1=2 q_s^2/\Gamma$
\cite{deGennes}.

The time evolution of the fluid density and velocity are respectively governed
by the continuity equation
$\partial_t \rho = -\partial_\alpha(\rho v_\alpha)$, and
the following Navier-Stokes equation:
\begin{eqnarray}
\partial_t v_\alpha +\rho \,v_\beta \partial_\beta v_\alpha
&=& \partial_\beta \Pi_{\alpha \beta}+ \eta\, \partial_\beta [ \partial_\alpha v_\beta +\; \partial_\beta v_\alpha].
\label{NSE}
\end{eqnarray}
This emerges from the Chapman-Enskog expansion
of the lattice Boltzmann (LB) equations 
that we solve numerically. A further term $\eta(1+3\frac{\partial P_0}{\partial\rho} )\partial_\mu v_\mu \delta_{\alpha \beta}$ that formally appears
in this expansion is negligible under the slow flows considered here
for which the the fluid motion is almost incompressible~\cite{Denniston:2001}.
$\eta$ is an isotropic background viscosity which is set to $\eta=0.6666$ in LB units (these are discussed below, see~\cite{Henrich:2011a,Henrich:2010b}).
The thermodynamic stress tensor reads explicitly
\begin{eqnarray}
\Pi_{\alpha \beta}&=&P_0 \delta_{\alpha\beta}
-\xi H_{\alpha \gamma}\left(Q_{\gamma \beta} +\frac{1}{3} \delta_{\gamma \beta}\right)\nonumber\\
&-&\xi \left(Q_{\alpha \gamma} +\frac{1}{3} \delta_{\alpha \gamma}\right) H_{\gamma \beta} + Q_{\alpha \gamma}H_{\gamma \beta}-H_{\alpha \gamma} Q_{\gamma \beta} \nonumber\\
&+&2 \xi  \left(Q_{\alpha \beta} +\frac{1}{3} \delta_{\alpha \beta}\right) Q_{\gamma \nu} H_{\gamma \nu}
- \partial_\alpha Q_{\gamma \nu} \frac{\delta{\cal F}}{\delta \partial_{\beta} Q_{\gamma \nu}}\nonumber\\
\label{Pi}
\end{eqnarray}
and is responsible for strong non-Newtonian flow effects.
In the isotropic state ${\bf Q}\equiv 0$ and Eq.\ref{Pi} reduces to the
scalar pressure as would appear in Eq.~\ref{NSE} for a Newtonian fluid.

We next define a 
dimensionless number that describes the deformation
of the director field under flow. This so-called Ericksen number
is given by 
\begin{equation}
Er=\frac{\eta v l}{K}
\end{equation}
with $\eta$ and $K$ defined previously, and 
$v$ and $l$ a typical velocity and length scale. 
In the present work $l=p_0/2=\pi/q_0$ was used as this is the approximate size of
the BP unit cell. Likewise, $v$ was taken to be the velocity difference
across one unit cell, i.e. $v=\dot{\gamma} l$. \rev{The elastic constants were $K=0.02$ (BPII) and $K=0.04$ (BPI), respectively.}
